---
title: "main"
author: "Constantin Goeldel"
date: "`r Sys.Date()`"
output: html_document
---

# The epimutation profile of gbmGenes

This script analyses patterns in epimutation rates in gene-body-methylated genes. The epimutation rates are estimated using [AlphaBeta](https://github.com/jlab-code/AlphaBeta)

## Install dependencies
```{r include=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE ))
    install.install.packages("BiocManager")
BiocManager::install(version = "3.16")

BiocManager::install("AlphaBeta")
BiocManager::install("data.table")
BiocManager::install("dplyr")
```

## Estimate epimutation rate

```{r}
nodeFile = "nodelist.fn"
edgeFile = "edgelist.fn"
setwd("/mnt/extStorage/constantin/epimutation-profile")
run.alphabeta.new(nodelist=nodeFile,
                  edelist=edgeFile,
                  name="epimutation_rate_estimation",
                  input.dir="/mnt/extStorage/constantin/epimutation-profile/",
                  output.dir="/mnt/extStorage/constantin/epimutation-profile/")
# estimate_epimutation(nodeFile,edgeFile, "epimutation_rate_estimation", "/mnt/extStorage/constantin/epimutation-profile", "/mnt/extStorage/constantin/epimutation-profile" ) 
```


## Building Pedigree files of MA lineages
```{r include=FALSE}
nodeFile = "./nodelist.fn"
edgeFile = "./edgelist.fn"

output = AlphaBeta::buildPedigree(nodelist = nodeFile, edgelist = edgeFile, cytosine = "CG", posteriorMaxFilter= 0.99 )

pedigree = output$Pdata 
```

##  Visualize pedigree
```{r}
AlphaBeta::plotPedigree(nodelist = nodeFile, edgelist = edgeFile, sampling.design = "progenitor.endpoint",
output.dir = out.dir, plot.width = 5, plot.height = 5, aspect.ratio = 1, vertex.size = 6,
vertex.label = FALSE)
```

## Divergence time vs methylome divergence

```{r}
dt = pedigree[, 2] + pedigree[, 3] - 2 * pedigree[, 1]
plot(dt, pedigree[, "D.value"], ylab = "Divergence value", xlab = expression(paste(Delta,
" t")))
```

## Epimutation rate estimation 

```{r}
# Initial proportion of unmethylated cytosines
p0uu_in = output$tmpp0

output.data.dir <- paste0(getwd())
estimates = AlphaBeta::ABneutral(pedigree.data = pedigree, p0uu = p0uu_in, eqp = p0uu_in, eqp.weight = 1,
Nstarts = 50, out.dir = paste0(getwd()), out.name = "ABneutral_CG_global_estimates")
summary(estimates)
```

## Analysis




